Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-07 arxiv/proof_examples.tex
finish section 5

%\subsubsection{Proof for Sample Imbalance}
We now give the proof of Proposition \ref{cor_MTL_loss}. Similar to the proof of Theorem \ref{thm_many_tasks}, our proof is divided into the following steps: 
\begin{itemize}
\item simplifying the optimization objective using concentration estimates;

\item finding the minimizer of the optimization objective;

\item deriving a bias-variance decomposition of the prediction loss $L(\hat \beta_2^{\MTL})$;

\item calculating the bias term and the variance term using Theorem \ref{thm_main_RMT}. 
\end{itemize}

@@

\paragraph{Simplifying the optimization objective.} 

@@

 \paragraph{Minimizer of the optimization objective.} We show that for the isotropic model, the minimizer of $g(A)$, denoted by $\hat A$, is very close to 1 under the condition \eqref{choiceofpara0}. This will be used to calculate the predication loss $\te(\hat{\beta}_2^{\MTL})$ loss.
%Our main observation is that the optimum of $\val(v)$ is approximately $1$.This allows us to simplify the bias and variance bounds significantly.
%Furthermore, under certain choice of parameters, we can simplify  
%More precisely, we have the following claim.
%Later on, we will combine it with Lemma \ref{prop_model_shift_tight} together to analyze the bias-variance tradeoff.

@@

\begin{proof}
Corresponding to equation \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}. \label{op_uplow}
\end{align}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(1)^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we used equation \eqref{choiceofpara0}. Hence using Claim \ref{claim_largedev1} and equations \eqref{eq_h1}-\eqref{eq_hA}, we obtain that with high probability,
\begin{align}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1) \label{gA>g1}
\end{align}
under condition \eqref{choiceofpara0}.
This gives that
%\be\label{rough A-1}
$|\hat A-1|\le \delta$ for any small constant $\delta>0$.

@@

Now we are ready to prove equation \eqref{hatw_add1}. First using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le A\le 1+\delta$,
$$h'(A)\ge \frac{2(A-1)\kappa^2 - d^2}{A^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+e_2}\le A\le 1+\delta$,
\begin{align*}
h(A)-h(1)\ge h(A)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^A h'(x)\dd x \gtrsim p\kappa^2 \cdot |A-1|^2  \gtrsim p\frac{d^2}{\kappa^2} + p^{-1/2+2e_2}\cdot p\kappa^2.
\end{align*}
Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
$$ h(A)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under equation \eqref{choiceofpara0}, which implies the inequality \eqref{gA>g1}. Second using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le A \le 1-d^2/\kappa^2$,
$$- h'(A)\ge  [2(1-A)\kappa^2 - Ad^2] \cdot p^{-1} \tr\left[\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-\frac12c_\infty}\le A\le 1-\delta$, then 
$$ h(A)-h(1) \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under equation \eqref{choiceofpara0}, which implies the inequality \eqref{gA>g1}. In sum, we obtain that the minimizer $\hat A$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Calculating the prediction loss.} Now we calculate the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat A=\hat A_1/\hat A_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left|\Sigma_2^{1/2} \left( \hat B \hat A_2 - \beta_2\right)\right| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat A\beta_1-\hat A^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat A   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2,\label{Lbeta_HPS}
\end{align}
where we denoted
$$\hat \Sigma:= \hat A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat A$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2} \nonumber \\
&	=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(2)})^\top}+ \sigma^2\cdot \hat A^2 \bigtr{ X^{(1)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}&	=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

Now we bound the left-hand side of \eqref{simple1} as

@@

% \todo{switch the $a_3$ and $a_4$}
\paragraph{Calculating the variance and bias terms.} Finally, we calculate the variance term $\sigma^2  \bigtr{\Sigma(1)^{-1}  }$ and the bias term $d^2\cdot {p}^{-1}\tr\left[\Sigma(1)^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right]$ using Theorem \ref{thm_main_RMT}, which will conclude Proposition \ref{cor_MTL_loss}.
%For the isotropic model in Section \ref{sec_similarity}, we actually have an easier and sharper bound than Lemma \ref{thm_model_shift}. %as follows.
%\begin{lemma}\label{prop_model_shift_tight}
%		In the setting of Lemma \ref{thm_model_shift}, assume that $\Sigma_1 =\id$,
%		$\beta_2$ is a random vector that has i.i.d. entries with mean $0$, variance $\kappa^2$ and finite moments up to any order, and $(\beta_1-\beta_2)$ is a random vector that is independent of $\beta_2$ and has i.i.d. entries with mean $0$, variance $d^2$ and finite moments up to any order. Denote
%		$\Delta^\star_{\bias} := \bigbrace{(1 - \hat{v})^2 \kappa^2 + d^2} \bigtr{\Pi(\hat v)}$, where $\Pi$ was defined in equation \eqref{defnpihat}.	Then we have with high probability,
%		\begin{align*}
%			\te(\hat{\beta}_t^{\MTL}) \le \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \ge  (\al_+^2(\rho_1)+\oo(1))\cdot  \Delta^\star_{\bias} \ , \\
%			\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \le (\al_-^2(\rho_1)-\oo(1))\cdot  \Delta^\star_{\bias}\ ,
%		\end{align*}
%		where $\Delta_{\vari}$ was defined in equation \eqref{Deltavarv}.
%\end{lemma}

>>>>

%\subsubsection{Proof for Sample Imbalance}
We provide the proof of Proposition \ref{cor_MTL_loss}. Similar to the proof of Theorem \ref{thm_many_tasks}, our proof is divided into the following parts:
\begin{itemize}
\item First, we reduce the optimization objective to another function $g(x)$ that only depends on the sample covariance matrices.
\item Second, we show that the global minimizer of $g(x)$ is approximate $1$.
\item Finally, we use the above result and Theorem \ref{thm_main_RMT} to prove the proposition.
\end{itemize}

@@

\paragraph{Part $1$: the reduction.}

@@

\paragraph{Part $2$: approximating the global minimum.}
	Let $\hat{A}$ denote the minimizer of $g(A)$.\HZ{change $\hat{A}$ to say $\hat{x}$.}
	We show that in the setting of Corollary \ref{cor_MTL_loss}, $\hat A$ is close to 1.% This will be used to calculate the predication loss $\te(\hat{\beta}_2^{\MTL})$ loss.
%Our main observation is that the optimum of $\val(v)$ is approximately $1$.This allows us to simplify the bias and variance bounds significantly.
%Furthermore, under certain choice of parameters, we can simplify  
%More precisely, we have the following claim.
%Later on, we will combine it with Lemma \ref{prop_model_shift_tight} together to analyze the bias-variance tradeoff.

@@

\begin{proof}
Corresponding to equation \eqref{EgA}, we define the function
\begin{align*}
h(A):= [(A-1)^2\kappa^2 + (A^2+1)d^2/2] \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)}\right].
\end{align*}
Using Fact \ref{lem_minv} (iii), we can obtain that with high probability,
\begin{align} 
&\left( A^2 \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^{-1} \right)^{-1} \nonumber\\
&\preceq (X^{(1)})^\top X^{(1)} \Sigma(A)^{-1}(X^{(2)})^\top X^{(2)} =\left( A^2 [(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1} \right)^{-1} \nonumber\\
&\preceq \left( A^2 \left[(\sqrt{n_2}+\sqrt{p})^{2}+n_2p^{-c_\varphi}\right]^{-1} +\left[(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}\right]^{-1} \right)^{-1}. \label{op_uplow}
\end{align}
Then we get that
\be\label{eq_h1}h(1)= d^2 \cdot p^{-1} \tr\left[ (X^{(1)})^\top X^{(1)} \Sigma(1)^{-1}(X^{(2)})^\top X^{(2)}\right]\lesssim pd^2\left(\rho_1^{-1} + \rho_2^{-1}\right)^{-1}\lesssim pd^2.\ee
where we used that $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$. On the other hand, if $|A-1|\ge \delta$ for a small constant $\delta>0$, then we have 
\begin{align}\label{eq_hA}
h(A)\gtrsim (1+A^2)\kappa^2 \cdot p\left(\rho_1^{-1} + A^2 \rho_2^{-1}\right)^{-1} \gtrsim p\kappa^2 ,
\end{align}
where in the first step we used that $|A-1|^2\gtrsim 1+A^2$ for $|A-1|\ge \delta$, $1+\tau\le \rho_1\le \tau^{-1}$ and $1+\tau\le \rho_2\le \tau^{-1}$, and in the second step we use equation \eqref{choiceofpara0} \HZ{I'm not sure which condition we use? put it here}. Hence using Claim \ref{claim_largedev1} and equations \eqref{eq_h1}-\eqref{eq_hA}, we obtain that with high probability,
\begin{align}
g(A)&\ge  h(A) +\sigma^2(n_1+n_2-p) -p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \nonumber\\
&\ge h(1) +\sigma^2(n_1+n_2-p) + p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right) \ge g(1), \label{gA>g1}
\end{align}
under condition \eqref{choiceofpara0} \HZ{again, I don't see which condition we use here?}.
This gives that
%\be\label{rough A-1}
$|\hat A-1|\le \delta$ for any small constant $\delta>0$.

@@

Now we are ready to prove equation \eqref{hatw_add1}. First using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le A\le 1+\delta$,
$$h'(A)\ge \frac{2(A-1)\kappa^2 - d^2}{A^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + A^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+e_2}\le A\le 1+\delta$,
\begin{align*}
h(A)-h(1)\ge h(A)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^A h'(x)\dd x \gtrsim p\kappa^2 \cdot |A-1|^2  \gtrsim p\frac{d^2}{\kappa^2} + p^{-1/2+2e_2}\cdot p\kappa^2.
\end{align*}
Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
$$ h(A)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right),$$
where we use the assumption that $\sigma^2$ and $d^2$ are both at most $\OO(\kappa^2)$, which implies the inequality \eqref{gA>g1}. Second using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le A \le 1-d^2/\kappa^2$,
$$- h'(A)\ge  [2(1-A)\kappa^2 - Ad^2] \cdot p^{-1} \tr\left[\left(A^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |A-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-\frac12c_\infty}\le A\le 1-\delta$, then 
$$ h(A)-h(1) \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right)$$
under equation \eqref{choiceofpara0} \HZ{I don't see which condition we want to use here?}, which implies the inequality \eqref{gA>g1}. In sum, we obtain that the minimizer $\hat A$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Part $3$: Applying Theorem \ref{thm_main_RMT}.}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat A=\hat A_1/\hat A_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left|\Sigma_2^{1/2} \left( \hat B \hat A_2 - \beta_2\right)\right| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat A\beta_1-\hat A^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat A   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2,\label{Lbeta_HPS}
\end{align}
where we denoted
$$\hat \Sigma:= \hat A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat A$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2} \nonumber \\
&	=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(2)})^\top}+ \sigma^2\cdot \hat A^2 \bigtr{ X^{(1)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}&	=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

Now we bound the left-hand side of \eqref{simple1} as \HZ{replace $\Sigma(1)$ with $\hat{\Sigma}$ throughout}

@@

% \todo{switch the $a_3$ and $a_4$}
Finally, we are ready to present the proof of Proposition \ref{cor_MTL_loss}.
We derive the variance term $\sigma^2  \bigtr{\Sigma(1)^{-1}  }$ and the bias term $d^2\cdot {p}^{-1}\tr\left[\Sigma(1)^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right]$ using Theorem \ref{thm_main_RMT}.
%For the isotropic model in Section \ref{sec_similarity}, we actually have an easier and sharper bound than Lemma \ref{thm_model_shift}. %as follows.
%\begin{lemma}\label{prop_model_shift_tight}
%		In the setting of Lemma \ref{thm_model_shift}, assume that $\Sigma_1 =\id$,
%		$\beta_2$ is a random vector that has i.i.d. entries with mean $0$, variance $\kappa^2$ and finite moments up to any order, and $(\beta_1-\beta_2)$ is a random vector that is independent of $\beta_2$ and has i.i.d. entries with mean $0$, variance $d^2$ and finite moments up to any order. Denote
%		$\Delta^\star_{\bias} := \bigbrace{(1 - \hat{v})^2 \kappa^2 + d^2} \bigtr{\Pi(\hat v)}$, where $\Pi$ was defined in equation \eqref{defnpihat}.	Then we have with high probability,
%		\begin{align*}
%			\te(\hat{\beta}_t^{\MTL}) \le \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \ge  (\al_+^2(\rho_1)+\oo(1))\cdot  \Delta^\star_{\bias} \ , \\
%			\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})\quad \text{ when: }\ \  & \Delta_{\vari} \le (\al_-^2(\rho_1)-\oo(1))\cdot  \Delta^\star_{\bias}\ ,
%		\end{align*}
%		where $\Delta_{\vari}$ was defined in equation \eqref{Deltavarv}.
%\end{lemma}
